Manuscript 1 code

### 2024-10-10 manuscript_1_code.Rmd
Updated figs, created spatial synchrony figs/analyses

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
library(dplyr)
library(janitor)
```

@@

## Figure formatting
```{r figure formatting info}
mytheme <- theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_rect(fill = "white"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))

@@

macro_species_cols <- macroalgae[,-which(names(macroalgae) %in% meta_cols)]
```

@@

# Figure 1

@@

## plot after aes changes
```{r Figure 1 v2}
plot_list <- lapply(levels(figure_1_data$habitat), function(hab) {
  ggplot(data = subset(figure_1_data, habitat == hab)) +
    geom_point(aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
    geom_errorbar(aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group), 
                  width = 0, size = 1.3) +
    geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
    annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
    geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
    geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
    scale_colour_manual(values = functional_group_colours) +
    scale_x_continuous(breaks = unique(figure_1_data$year)) +
    labs(y = "Percent Cover",
         x = "Year",
         title = hab) +  # Add habitat as title
    theme_bw() +
    mytheme
})

figure_1 <- ggarrange(plotlist = plot_list, 
                      ncol = 2, nrow = 2, 
                      labels = c("a.", "b.", "c.", "d."),
                      common.legend = TRUE,
                      legend = "bottom", 
                      font.label = list(size = 20, color = "black", face = "plain"))
# 
# figure_1_caption <- "Figure 1. Benthic coverage of four major space holders in Moorea across four habitats (Fringing, Backreef, Forereef at 10m depth, Forereef  at 17m depth). Annual fluctuations in percent cover of macroalgae, coral, turf, and crustose coralline algae (CCA) are displayed as points for the average cover for each benthic group across sites at each habitat with lines showing standard error. Dotted lines show disturbance events for Cyclone Oli (2010) in black, and the 2019 bleaching event in red, and the gray horizontal line shows crown of thorns (COTs) outbreak which occurred from 2007-2010."
# 
# figure_1_with_caption <- annotate_figure(figure_1,
#                 bottom = text_grob(figure_1_caption, hjust = 0.5, vjust = 1, size = 10))

#ggsave(filename = "output/figure_1v2.png", figure_1, height = 10, width = 14)
```

# Figure 2

@@

```{r}
macroalgal_cover <- cover_df[,c("year", "habitat", "site", "location", "macroalgae")]

alpha_diversity_quad_macro_with_cover <- left_join(alpha_diversity_quad_macro, macroalgal_cover,by = join_by(location, year, site, habitat)) %>% 
  mutate(macroalgal_prop_cover = macroalgae / 100) %>% 
  select(-macroalgae)
```

@@

### plot
```{r figure 2a}
(figure_2a <- ggplot() +
  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
             position = position_dodge(0.4)) +
  geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se,
                                             colour = habitat), width = 0, size = 1.3, position = position_dodge(0.4)) +
  scale_colour_manual(values = habitat_colours) +
  # scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(quad_richness_df$year)) +
   labs(y = "Macroalgal Richness",
        x = "Year",
        title = "a. Quadrat-level Richness") +
  theme_bw() + 
  mytheme +
   theme(axis.text.x = element_text(size=20),
         axis.title.x = element_text(size=22),
         axis.title.y = element_text(size=22),
      axis.text.y = element_text(size=20),
      plot.title = element_text(size=24),
      legend.text = element_text(size=20),
      legend.title = element_blank()))

@@

### plot
```{r plot figure 2b}
(figure_2b <- ggplot() +
  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
             position = position_dodge(0.4)) +
#  geom_jitter(data = alpha_diversity_site_macro, 
#  aes(x = year, y = richness, colour = habitat, alpha = 0.05), size = 2, width = 0.4, height = 0) +
  geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0,  position = position_dodge(0.4)) +
  scale_colour_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 12, yend = 12, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(site_richness_df$year)) +
   labs(y = "",
        x = "Year",
        title = "b. Site-level Richness") +
  mytheme +
   theme(axis.text.x = element_text(size=20),
         axis.title.x = element_text(size=22),
      axis.text.y = element_text(size=20),
      plot.title = element_text(size=24),
      legend.text = element_text(size=20),
      legend.title = element_blank(),
      ))

@@

## Figure 2 Full Plot
```{r ggpubr fig2, fig.height = 10, fig.width=22}
fig2_plot_list <- list(figure_2a, figure_2b)
(figure_2 <- ggarrange(plotlist = fig2_plot_list, 
                      ncol = 2, nrow = 1, 
                      common.legend = TRUE,
                      legend = "bottom", 
                      font.label = list(size = 20, color = "black", face = "plain")))

# ggsave(filename = "output/figure_2v2.png", figure_2, height = 10, width = 22)
```

# Figure 3

## set up site-level data for model
```{r site cover by diversity data setup}
# alpha_diversity_site_macro_with_cover <- alpha_diversity_site_macro %>% 
#   dplyr::mutate(percent_cover = rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])) %>% 
#   mutate(prop_cover = percent_cover / 100) %>% 
#   select(-percent_cover) %>% 
#   mutate(cover_trans = sv_trans(prop_cover))

#there is no reasonable reason why the above chunk doesn't work, but it doesn't, so we did it this way
alpha_diversity_site_macro$percent_cover <- rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])

alpha_diversity_site_macro_with_cover <- alpha_diversity_site_macro %>% 
  dplyr::mutate(prop_cover = percent_cover / 100) %>%
  select(-percent_cover)

alpha_diversity_site_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_site_macro_with_cover$prop_cover)
```

@@

### quad
```{r figure 3a richness quad model}
alpha_diversity_quad_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_quad_macro_with_cover$macroalgal_prop_cover)

cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_with_cover)
summary(cover_mod)
Anova(cover_mod)
#                     Chisq Df Pr(>Chisq)    
# richness         30916.73  1  < 2.2e-16 ***
# habitat            141.78  3  < 2.2e-16 ***
# richness:habitat   294.54  3  < 2.2e-16 ***
hist(residuals(cover_mod)) # looks fine
plot(residuals(cover_mod) ~ fitted(cover_mod)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
r.squaredGLMM(cover_mod) # 0.631, 0.650
emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
```

## figure 3a plot setup
```{r figure 3a setup}
diversity_ranges_quad <- alpha_diversity_quad_macro_with_cover %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness),
            max_richness = max(richness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

@@

# alpha_diversity_quad_macro_with_cover$habitat <- factor(alpha_diversity_quad_macro_with_cover$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
# 
# 
# filtered_cover_effects$habitat <- factor(filtered_cover_effects$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

@@

```{r figure 3a plot}
(figure_3a <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad_macro_with_cover,
             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
  geom_line(data = filtered_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.5) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover\n", x = "", title = "3a.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=22),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
# ggsave(filename = "output/figure_3a.png", figure_3a, height = 10, width = 14)
```

@@

#merge functional groups with 
macro_functional_groups <- merge(macro_long_data, algae_fg, all.x = TRUE)

@@

### model
```{r functional richness cover mod quad}
#merge Noam's data with functional group summary 
functional_richness_quad_cover <- merge(alpha_diversity_quad_macro_with_cover, fg_summary_wide,
                                        by = c("location", "year", "site", "habitat"))

@@

r.squaredGLMM(cover_mod_fg) # 0.64, 0.65
emtrends(cover_mod_fg, pairwise ~ habitat, var = "fg_richness_quad") 
#all habitats are different from forereef 17

@@

cover_emm <- emmip(cover_mod_fg,
                        habitat ~ fg_richness_quad,
                        at = list(fg_richness_quad = seq(0,10,1)), plotit = F, CIs = T)

filtered_cover_effects <- cover_emm %>%
  left_join(diversity_ranges_quad_fg, by = c("habitat")) %>%
  filter(fg_richness_quad >= min_fg_richness_quad & fg_richness_quad <= max_fg_richness_quad)

@@

```{r figure 3b plot}
(figure_3b <- 
  ggplot() +
  geom_point(data = functional_richness_quad_cover,
             aes(x = fg_richness_quad, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
  geom_line(data = filtered_cover_effects, aes(x = fg_richness_quad, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects, aes(x = fg_richness_quad, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "", title = "3b.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=22),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

@@

### synchrony set up
```{r calculate syncrony}
# make er long
macro_long_data <- pivot_longer(
  data = macroalgae,
  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
  names_to = "taxa",
  values_to = "percent_cover") %>% 
  mutate(prop_cover = percent_cover/100) %>% 
  select(-percent_cover)

@@

diversity_stability <- alpha_diversity_quad_macro_with_cover %>%
  dplyr::group_by(location, habitat, site) %>%
  dplyr::summarise(
            richness_mean = mean(richness), 
            shannon_mean = mean(shannon), 
            cover_mean = mean(macroalgal_prop_cover),
            evenness_mean = mean(evenness, na.rm = T),
            cover_cv = CV(macroalgal_prop_cover),
            cover_stability = 1/CV(macroalgal_prop_cover))

diversity_stability_synchrony <- left_join(diversity_stability, lm_synchrony, by = join_by(location))
diversity_stability_synchrony$synchrony_trans <- sv_trans(diversity_stability_synchrony$synchrony)
# diversity_stability_synchrony <- na.omit(diversity_stability_synchrony) # remove quads that never had algae --> NaN calcs

@@

r.squaredGLMM(synch_cov_mod) # error??
emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats
```

@@

```{r synchrony cover plot}
(figure_3c <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.3) + 
  geom_line(data = filtered_synch_cov_effects, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_synch_cov_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "", y = "", title = "3c.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=22),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

@@

diversity_stability_site <- alpha_diversity_site_macro_with_cover %>%
  dplyr::group_by(site_habitat, habitat, site) %>%
  dplyr::summarise(
            richness_mean = mean(richness), 
            shannon_mean = mean(shannon), 
            cover_mean = mean(prop_cover),
            evenness_mean = mean(evenness),
            cover_cv = CV(prop_cover),
            cover_stability = 1/CV(prop_cover))

diversity_stability_synchrony_site <- left_join(diversity_stability_site, lm_synchrony_site, by = join_by(site_habitat))
diversity_stability_synchrony_site$synchrony_trans <- sv_trans(diversity_stability_synchrony_site$synchrony)
# diversity_stability_synchrony_site <- na.omit(diversity_stability_synchrony_site) # remove quads that never had algae --> NaN calcs
# write_csv(diversity_stability_synchrony_site, here("data", "diversity_stability_synchrony_site.csv"))

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
library(dplyr)
library(janitor)
library(knitr)
```

@@

## Figure formatting
```{r figure formatting info}
mytheme <- theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=24), 
        plot.title = element_text(size = 24),
        panel.background = element_rect(fill = "white"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))

@@

macro_species_cols <- macroalgae[,-which(names(macroalgae) %in% meta_cols)]

# make er long
macro_long_data <- pivot_longer(
  data = macroalgae,
  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
  names_to = "taxa",
  values_to = "percent_cover") %>% 
  mutate(prop_cover = percent_cover/100) %>% 
  select(-percent_cover)

# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))
```

@@

# Figure 1 - benthic cover trends among habitats

@@

## plot after aes changes
```{r Figure 1 v2, fig.height = 10, fig.width=14}
plot_list <- lapply(levels(figure_1_data$habitat), function(hab) {
  ggplot(data = subset(figure_1_data, habitat == hab)) +
    geom_point(aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
    geom_errorbar(aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group), 
                  width = 0, size = 1.3) +
    geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
    annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
    geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
    geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
    scale_colour_manual(values = functional_group_colours) +
    scale_x_continuous(breaks = unique(figure_1_data$year)) +
    labs(y = "Percent Cover",
         x = "Year",
         title = hab) +  # Add habitat as title
    theme_bw() +
    mytheme
})

(figure_1 <- ggarrange(plotlist = plot_list, 
                      ncol = 2, nrow = 2, 
                      labels = c("a.", "b.", "c.", "d."),
                      common.legend = TRUE,
                      legend = "bottom", 
                      font.label = list(size = 24, color = "black", face = "plain")))
# 
# figure_1_caption <- "Figure 1. Benthic coverage of four major space holders in Moorea across four habitats (Fringing, Backreef, Forereef at 10m depth, Forereef  at 17m depth). Annual fluctuations in percent cover of macroalgae, coral, turf, and crustose coralline algae (CCA) are displayed as points for the average cover for each benthic group across sites at each habitat with lines showing standard error. Dotted lines show disturbance events for Cyclone Oli (2010) in black, and the 2019 bleaching event in red, and the gray horizontal line shows crown of thorns (COTs) outbreak which occurred from 2007-2010."
# 
# figure_1_with_caption <- annotate_figure(figure_1,
#                 bottom = text_grob(figure_1_caption, hjust = 0.5, vjust = 1, size = 10))

# ggsave(filename = "output/figure_1v3.png", figure_1, height = 10, width = 14)
```

# Figure 2 - macroalgal richness trends among habitats

@@

```{r add cover to quad df}
macroalgal_cover_quad <- cover_df[,c("year", "habitat", "site", "location", "macroalgae")]

alpha_diversity_quad_macro <-
  left_join(alpha_diversity_quad_macro, macroalgal_cover_quad,by = join_by(location, year, site, habitat)) %>% 
  mutate(macroalgal_prop_cover = macroalgae / 100) %>% 
  select(-macroalgae)
```

@@

### plot
```{r figure 2a}
(figure_2a <- ggplot() +
  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
             position = position_dodge(0.4)) +
  geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se,
                                             colour = habitat), width = 0, size = 1.3, position = position_dodge(0.4)) +
  scale_colour_manual(values = habitat_colours) +
  # scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(quad_richness_df$year)) +
   labs(y = "Macroalgal Richness",
        x = "Year",
        title = "Quadrat-level Richness") +
  theme_bw() + 
  mytheme +
   theme(axis.text.x = element_text(size=20),
         axis.title.x = element_text(size=22),
         axis.title.y = element_text(size=22),
      axis.text.y = element_text(size=20),
      plot.title = element_text(size=24),
      legend.text = element_text(size=20),
      legend.title = element_blank()))

@@

### plot
```{r plot figure 2b}
(figure_2b <- ggplot() +
  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4,
             position = position_dodge(0.4)) +
#  geom_jitter(data = alpha_diversity_site_macro, 
#  aes(x = year, y = richness, colour = habitat, alpha = 0.05), size = 2, width = 0.4, height = 0) +
  geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0,  position = position_dodge(0.4)) +
  scale_colour_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 12, yend = 12, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(site_richness_df$year)) +
   labs(y = "",
        x = "Year",
        title = "Site-level Richness") +
  mytheme +
   theme(axis.text.x = element_text(size=20),
         axis.title.x = element_text(size=22),
      axis.text.y = element_text(size=20),
      plot.title = element_text(size=24),
      legend.text = element_text(size=20),
      legend.title = element_blank(),
      ))

@@

## Figure 2 Full Plot
```{r ggpubr fig2, fig.height = 10, fig.width=22}
fig2_plot_list <- list(figure_2a, figure_2b)
(figure_2 <- ggarrange(plotlist = fig2_plot_list, 
                      ncol = 2, nrow = 1, 
                      labels = c("a.", "b."),
                      common.legend = TRUE,
                      legend = "bottom", 
                      font.label = list(size = 24, color = "black", face = "plain")))

# ggsave(filename = "output/figure_2v3.png", figure_2, height = 10, width = 22)
```

# Figure 3 - ecosystem function

@@

### quad
```{r figure 3a richness quad model}
alpha_diversity_quad_macro$cover_trans <- sv_trans(alpha_diversity_quad_macro$macroalgal_prop_cover)

cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro)
summary(cover_mod)
Anova(cover_mod)
#                     Chisq Df Pr(>Chisq)    
# richness         30916.73  1  < 2.2e-16 ***
# habitat            141.78  3  < 2.2e-16 ***
# richness:habitat   294.54  3  < 2.2e-16 ***
hist(residuals(cover_mod)) # looks fine
plot(residuals(cover_mod) ~ fitted(cover_mod)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
r.squaredGLMM(cover_mod) # 0.631, 0.650
emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef          0.34720 0.0241 Inf  14.409  <.0001
# Fringing - Forereef 10m      0.34865 0.0240 Inf  14.514  <.0001
# Fringing - Forereef 17m      0.38953 0.0250 Inf  15.599  <.0001
# Backreef - Forereef 10m      0.00144 0.0205 Inf   0.071  0.9999
# Backreef - Forereef 17m      0.04232 0.0214 Inf   1.979  0.1957
# Forereef 10m - Forereef 17m  0.04088 0.0209 Inf   1.952  0.2063
```

## figure 3a plot setup
```{r figure 3a setup}
diversity_ranges_quad <- alpha_diversity_quad_macro %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness),
            max_richness = max(richness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

@@

@@

```{r figure 3a plot}
(figure_3a <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad_macro,
             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
  geom_line(data = filtered_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.5) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover\n", x = "", title = "") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=22),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
# ggsave(filename = "output/figure_3a.png", figure_3a, height = 10, width = 14)
```

@@

#merge functional groups  
macro_functional_groups <- merge(macro_long_data, algae_fg, all.x = TRUE)

@@

### model
```{r functional richness cover mod quad}
#merge Noam's data with functional group summary 
functional_richness_quad_cover <- merge(alpha_diversity_quad_macro, fg_summary_wide,
                                        by = c("location", "year", "site", "habitat"))

@@

r.squaredGLMM(cover_mod_fg) # 0.64, 0.65
emtrends(cover_mod_fg, pairwise ~ habitat, var = "fg_richness_quad") 
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef            0.214 0.0266 Inf   8.051  <.0001
# Fringing - Forereef 10m        0.346 0.0261 Inf  13.252  <.0001
# Fringing - Forereef 17m        0.530 0.0262 Inf  20.244  <.0001
# Backreef - Forereef 10m        0.132 0.0231 Inf   5.708  <.0001
# Backreef - Forereef 17m        0.316 0.0232 Inf  13.617  <.0001
# Forereef 10m - Forereef 17m    0.184 0.0223 Inf   8.241  <.0001

@@

cover_emm_fg <- emmip(cover_mod_fg,
                        habitat ~ fg_richness_quad,
                        at = list(fg_richness_quad = seq(0,10,1)), plotit = F, CIs = T)

filtered_cover_effects_fg <- cover_emm_fg %>%
  left_join(diversity_ranges_quad_fg, by = c("habitat")) %>%
  filter(fg_richness_quad >= min_fg_richness_quad & fg_richness_quad <= max_fg_richness_quad)

@@

```{r figure 3b plot}
(figure_3b <- 
  ggplot() +
  geom_point(data = functional_richness_quad_cover,
             aes(x = fg_richness_quad, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
  geom_line(data = filtered_cover_effects_fg, aes(x = fg_richness_quad, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects_fg, aes(x = fg_richness_quad, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "", x = "", title = "") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=22),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

@@

### synchrony set up
```{r calculate syncrony}

@@

diversity_stability <- alpha_diversity_quad_macro %>%
  dplyr::group_by(location, habitat, site) %>%
  dplyr::summarise(
            richness_mean = mean(richness), 
            shannon_mean = mean(shannon), 
            cover_mean = mean(macroalgal_prop_cover),
            evenness_mean = mean(evenness, na.rm = T),
            cover_cv = CV(macroalgal_prop_cover),
            cover_stability = 1/CV(macroalgal_prop_cover))

diversity_stability_synchrony <- left_join(diversity_stability, lm_synchrony, by = join_by(location))
diversity_stability_synchrony$synchrony_trans <- sv_trans(diversity_stability_synchrony$synchrony)

@@

r.squaredGLMM(synch_cov_mod) # error??
emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats

# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            0.101 0.290 Inf   0.348  0.9856
# Fringing - Forereef 10m        0.383 0.365 Inf   1.048  0.7211
# Fringing - Forereef 17m        3.458 0.484 Inf   7.140  <.0001
# Backreef - Forereef 10m        0.282 0.378 Inf   0.746  0.8784
# Backreef - Forereef 17m        3.357 0.492 Inf   6.820  <.0001
# Forereef 10m - Forereef 17m    3.075 0.501 Inf   6.142  <.0001

@@

```{r synchrony cover plot}
(figure_3c <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.3) + 
  geom_line(data = filtered_synch_cov_effects, aes(x = synchrony, y = exp(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_synch_cov_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "", y = "", title = "") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=22),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

@@

diversity_stability_site <- alpha_diversity_site_macro %>%
  dplyr::group_by(site_habitat, habitat, site) %>%
  dplyr::summarise(
            richness_mean = mean(richness), 
            shannon_mean = mean(shannon), 
            cover_mean = mean(prop_cover),
            evenness_mean = mean(evenness),
            cover_cv = CV(prop_cover),
            cover_stability = 1/CV(prop_cover))

diversity_stability_synchrony_site <- left_join(diversity_stability_site, lm_synchrony_site, by = join_by(site_habitat))
diversity_stability_synchrony_site$synchrony_trans <- sv_trans(diversity_stability_synchrony_site$synchrony)

# write_csv(diversity_stability_synchrony_site, here("data", "diversity_stability_synchrony_site.csv"))
